Differential expression analysis: A pipeline to analyse time-course gene expression data

### 2019-03-05 scripts/02_differential_expression_analysis.Rmd
DOC more on the paper

```{r echo=FALSE, results="hide"}
# Run the DE analysis on the rest of the influenza strain.
contrasts = c("M.0-K.0",
	      "M.3-K.3",
	      "M.6-K.6",
	      "M.9-K.9",
	      "M.12-K.12",
	      "M.18-K.18",
	      "M.24-K.24",
	      "M.30-K.30",
	      "M.36-K.36",
	      "M.48-K.48",
	      "M.60-K.60",
	      "M.72-K.72",
	      "M.120-K.120",
	      "M.168-K.168")
K_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, meta, contrasts,
     use_voom_weights=FALSE)
contrasts = c("M.0-C.0",
	      "M.3-C.3",
	      "M.6-C.6",
	      "M.9-C.9",
	      "M.12-C.12",
	      "M.18-C.18",
	      "M.24-C.24",
	      "M.30-C.30",
	      "M.36-K.36",
	      "M.48-C.48",
	      "M.60-C.60",
	      "M.72-C.72",
	      "M.120-C.120",
	      "M.168-C.168")
C_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, meta, contrasts,
     use_voom_weights=FALSE)
contrasts = c("M.0-VH.0",
	      "M.3-VH.3",
	      "M.6-VH.6",
	      "M.9-VH.9",
	      "M.12-VH.12",
	      "M.18-VH.18",
	      "M.24-VH.24",
	      "M.30-VH.30",
	      "M.36-VH.36",
	      "M.48-VH.48",
	      "M.60-VH.60",
	      "M.72-VH.72",
	      "M.120-VH.120",
	      "M.168-VH.168")
VH_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, meta, contrasts,
     use_voom_weights=FALSE)
```

@@

@@

@@

pvalues = moanin::timecourse_differential_expression_analysis(
    data, meta, contrasts)
qvalues = apply(pvalues, 2, stats::p.adjust)

@@

```{r echo=FALSE, results="hide"}
# Rename stuff for saving.
outname = ".results/pvalues.txt"
colnames(pvalues) = unlist(
    lapply(colnames(pvalues), function(x){paste0(x, "-pval")}))

>>>>

```{r echo=FALSE, results="hide"}
# Hide those results
# Run the DE analysis on the rest of the influenza strain.
contrasts = c("M.0-K.0",
	      "M.3-K.3",
	      "M.6-K.6",
	      "M.9-K.9",
	      "M.12-K.12",
	      "M.18-K.18",
	      "M.24-K.24",
	      "M.30-K.30",
	      "M.36-K.36",
	      "M.48-K.48",
	      "M.60-K.60",
	      "M.72-K.72",
	      "M.120-K.120",
	      "M.168-K.168")
K_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, meta, contrasts,
     use_voom_weights=FALSE)
contrasts = c("M.0-C.0",
	      "M.3-C.3",
	      "M.6-C.6",
	      "M.9-C.9",
	      "M.12-C.12",
	      "M.18-C.18",
	      "M.24-C.24",
	      "M.30-C.30",
	      "M.36-K.36",
	      "M.48-C.48",
	      "M.60-C.60",
	      "M.72-C.72",
	      "M.120-C.120",
	      "M.168-C.168")
C_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, meta, contrasts,
     use_voom_weights=FALSE)
contrasts = c("M.0-VH.0",
	      "M.3-VH.3",
	      "M.6-VH.6",
	      "M.9-VH.9",
	      "M.12-VH.12",
	      "M.18-VH.18",
	      "M.24-VH.24",
	      "M.30-VH.30",
	      "M.36-VH.36",
	      "M.48-VH.48",
	      "M.60-VH.60",
	      "M.72-VH.72",
	      "M.120-VH.120",
	      "M.168-VH.168")
VH_weekly_de_analysis =  moanin::weekly_differential_expression_analysis(
     data, meta, contrasts,
     use_voom_weights=FALSE)
```

@@

The distribution of number of genes found differentially expressed by
considering each time-point independantly highlights the challenges of such
approach. Timepoint 6H and 18H of the Kawasaki strain have fewer numbers of
genes found differentially expressed than timepoints 3H and 9H: this is likely
due to biological or technical variances for specific genes at specific
timepoints.

As a summary, classic differential expression methods are appropriate for
unordered treatments, but fail to use the temporal structure of the data.

@@

To leverage this temporal structure, Storey et al [CITE FIXME] proposes to
model time-course micro-array with splines functions, and to use a log-ratio
likelihood test to detect differentially expressed genes. `moanin` extends
this idea by providing functions to perform time-course differential
expression analysis using `limma` contrasts.

@@

# The function takes the data (data.frame or named matrix), the meta data
# (data.frame containing a timepoint and group column, the first corresponding
# to the time-course information, the latter corresponding to the
# treatment).
pvalues = moanin::timecourse_differential_expression_analysis(
    data, meta, contrasts)
qvalues = apply(pvalues, 2, stats::p.adjust)

@@

```{r echo=FALSE, results="hide"}
# Rename stuff for saving
outname = ".results/pvalues.txt"
colnames(pvalues) = unlist(
    lapply(colnames(pvalues), function(x){paste0(x, "-pval")}))
